Length Gauge Theory of Surface Second Harmonic Generation

### 2014-11-18 paper-v-0.tex
results
first round of corrections to results

\documentclass[floatfix,prb,aps,superscriptaddress,11pt,preprint,letterpaper]{revtex4}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\allowdisplaybreaks[1]%biutiful equation breaker!!!
\usepackage{ulem}
\usepackage{subfigure}
%\input{/Users/bms/util/definitions}
\def\chon{black}
\input{definitions}
\usepackage{bm}
%%%%*****************%%%% fine hyperef 
\usepackage[backref,pdffitwindow,colorlinks,linkcolor={blue},citecolor={red}]{hyperref}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{showlabels}
%\usepackage{showkeys}

%%%
% notes:
% nonlocal or non-local
%%%

@@

Concerning the optical response of surfaces and interfaces,
Mendoza and Moch\'an\cite{mendozaPRB96} produced an early review of 
theoretical SHG spectra from surfaces. 
Arzate and Mendoza\cite{arzatePRB01} improved upon that work and
went into a detailed analysis of the different contributions to the 
SHG spectra of a surface and the nuanced relationship between bulk, 
surface, interband, intraband, $1\omega$ and $2\omega$ terms. 
Calculated surface SHG spectra using these schemes coincide with experimental
results in both the velocity gauge
\cite{mendozaPRB01,mendozaPRL98,reiningPRB94} and the length gauge approach
\cite{sharmaPRB03}. 
Perhaps of more relevance to this work, 
Mendoza \emph{et al}.\cite{mendozaPRB06} and 
Mej\'ia \emph{et al}.\cite{mejiaRMF04} developed a layer-by-layer analysis for 
the linear and nonlinear responses of semiconductor systems. This model
allows for obtaining results from selected regions within a slab or bulk,
including the surface.
More recently, 
Trolle \emph{et al}.\cite{trollePRB14} 
elaborate on a new theoretical model for surface SHG of 
mono-layer MoS$_{2}$. Their approach is based on the tight-binding model 
fitted to GW results and include excitonic effects from the 
Bethe-Salpeter exciton picture. However, this new model is not 
\emph{ab initio} and has no consideration for a layered approach.

@@

In this section we present a test case in order to check the
consistency of the approach presented above. To do this end,  we have chosen
a clean Si(001) surface  with a $2\times 1$ surface reconstruction.
The slab for such a surface could be chosen to be centrosymmetric 
by having the fron and the back surface with the same $2\times 1$
reconstruction, however we terminate with hydrogen one of the
surfaces, thus having it as an ideally bulk terminated Si
surface. Indeed, the H atoms simply saturate the dangling bonds of the
bulk-like Si atoms at the surface, as seen in Fig.~\ref{si2x1}. \textcolor{red}{details of the
 relaxation and some references}
\begin{figure}
\centering 
\includegraphics[scale=.8]{images/si2x1-crop}
\caption{(color on line) The slab shows a front clean Si(001)$2\times 1$ surface,
  while the back surface is ideally terminated Si bulk, where the  
  dangling bonds are H (small balls) saturated. This picture shows 12  
  Si atomic layers and one H atomic layer. 
\label{si2x1}} 
\end{figure} 
The idea for such an slab is that the cristalline symmetry of the
H terminated surface imposes that $\chi_{\mathrm{H}}^{xxx}=0$, while
for the $2\times 1$ surface there is not such symmetry restriction and
then, $\chi_{2\times 1}^{xxx}\ne 0$.
Therefore, due to this fact, calculating $\chi^{xxx}$ for the full-slab
or the half-salb that contains the $2\times 1$ surface,\cite{nota1}
 ought to give the same result, since the contribution from the H
 saturated  surface is zero any way. Then, for this slab one must check
 whether or not this relation is satisfied,
\begin{align}\label{hs}
\chi_{\mathrm{half-slab}}^{xxx}(-2\go;\go,\go) 
=
\chi_{\mathrm{full-slab}}^{xxx}(-2\go;\go,\go)
,
\end{align}
where
$\chi_{\mathrm{half-slab}}^{xxx}(-2\go;\go,\go)$ is calculated using
$C^\ell(z)=1$ un the upper half of the salb, that contains the
$2\times 1$ surface reconstruction, as seen in
Fig.~\ref{si2x1},
whereas $\chi_{\mathrm{full-slab}}^{xxx}(-2\go;\go,\go)$ is calculated using
$C^\ell(z)=1$ through the full slab.
Indeed, in what follows we show the results of such a comparison.

The self-consistent ground state and their Kohn-Sham states were
calculated in the frame-work of DFT-LDA,
with the use of the plane-wave ABINIT
code.\cite{abinit}
We used
Troullier-Martins
pseudopotentials,\cite{troullierPRB91} 
that are 
fully separable nonlocal pseudopotentials in the 
Kleinman-Bylander 
form;\cite{kleinmanPRL82}
the contribution of $\bfv^\nl$ to Eq.~\eqref{chis},
can be calculated, as explained in Appendix \ref{vesnl}; 
we use the DP code for this end.\cite{francesco}
The $2\times 1$ and the H surfaces are relaxed \textcolor{red}{Nicolas 
  some details}, where we find agreement with previous results for 
both surfaces.\cite{relax} For instance the Si-H bond distance is 1.48 
$\AA$ ??. The unit cell has two atoms, making the calculation
computational very heavy. The vacuum size is equivalent to one quarter
of the
size of the slab, thus avoiding spurious effects from the possible
tunneling of the wave function of the contiguous surfaces of the full
crystal formed by the repeated super cell scheme.\cite{mendozaPRB06}

Spin-orbit effects, local field effects, and the consequences of the
electron-hole
attraction\cite{leitsmannPRB05,trollePRB14} on the SHG process are neglected. 
Although all these effects are important for the optical response of a
semiconductor, 
their calculation is still an open question and a numerical challenge
that ought to be pursued. 
However this endeavor is beyond the scope of this paper. 
For a given size  
of the slab,  we find converged spectra
for all the quantities of interest in this
work, 
and the most important parameters are
an energy cut-off of 10 Ha,
a number 
of conduction bands equal to the number of valence 
bands,
and a set of 244 $\bfk$-points
is used for
the linear analytic tetrahedron method used in the evaluation
the Brillouin zone integrals for of Eq.~\eqref{chis}, where special care
was taken to examine the double resonances.\cite{nastosPRB05}
Double resonances occur if for a given frequency $\omega$
there can be resonant
transitions at both frequencies $\omega$ and 2$\omega$,
 that is, if there is a
region in the Brillouin zone such that
$\omega_{cv}(\bfk)=2 \omega_{c'v}(\bfk)$
For these $\bfk$ points the perturbation theory used to calculate
the spectrum breaks down, since there is real population excited,
which in a correct calculation must be taken into account. These
points introduce sharp spikes in the spectrum that can, in
principle, affect the low-frequency results, since the response at
frequencies below the band gap is computed from the Kramers-Kronig
relation. However, in agreement with Ref.~\onlinecite{nastosPRB05}
 we find here
that the double resonances affect the low-frequency results by less
than 2\%. 
Finally, all the spectra are calculated with an energy Gaussian smearing of  0.15 eV {\color{red}Chon??} 
\begin{figure}
\centering 
\includegraphics[scale=.8]{plots/fig1}
\caption{(color on line) 
$|\chi_{\mathrm{half-slab}}^{xxx}|$ vs $\hbar\omega$ for different slabs with 16, 24, 32 
and 40 atomic Si layers. The fron surface of the slab presents a 
$2\times 1$ reconstruction, whereas the back sufrace is ideally bulk 
terminated where the dangling bonds are H-saturated. 
\label{fig1}} 
\end{figure}
\begin{figure}
\centering 
\includegraphics[scale=.8]{plots/fig2}
\caption{(color on line) 
$\chi^{xxx}_{\mathrm{half-slab}}$
and 
$\chi^{xxx}_{\mathrm{full-slab}}$
vs $\hbar\omega$ for a slab of 32 
atomic Si layers plus one layer of H. 
\label{fig2}} 
\end{figure}

To obtain Eqs.~\eqref{tau.1}  and \eqref{tau.1n}, required in
Eq.~\eqref{chis},
we see that the
evaluation of the commutator $[\bfr,\bfv^\nl]$ is required as it is
outlined in Appendix \ref{appvnl}; second order derivatives
are required to do such evaluation, making the numerical
procedure very
time consuming. Indeed this time is much higher than the already
lengthy time needed for the calculation of the contribution of
$\bfv^\nl$, proportional to only first order derivatives, as seen in
Appendix \ref{vesnl}. 
On top of this, the memory required is also very demanding for both
$\bfv^\nl$ and $[\bfr,\bfv^\nl]$.
However, the contribution coming from 
$[\bfr,\bfv^\nl]$, turns out to be very small,\cite{valerie} and
therefore we neglect it in this work.

In Fig.~\ref{fig1} we show $|\chi_{\mathrm{half-slab}}^{xxx}|$
 for a
slab of 16, 24, 32, and 40 Si atomic layers. We see that 
$|\chi_{\mathrm{half-slab}}^{xxx}|$ for
16 and 24
layers are different from each other, and different from that of 32
and 40 layers,
that are very similar to each other.
As explained above,
the calculation of the contribution given by
$\bfv^\nl$ is computational very demanding. A good
compromise between the accuracy in the convergence of
$\chi^{xxx}_{\mathrm{half-slab}}$,
 as a function of the number
of atomic layers in the slab, and the computational burden, is to take
the slab with 32 Si atomic layers, as representative of our system
in question.
In Fig.~\ref{fig2}
we compare 
$\chi^{xxx}_{\mathrm{half-slab}}$  
vs. 
$\chi^{xxx}_{\mathrm{full-slab}}$, 
for the four different possibilities of including or not including the
effects of $\bfv^\nl$ and having or not having a scissors correction
$\hbar\sigma$.
We see that for all four instances the comparison between
$\chi^{xxx}_{\mathrm{half-slab}}$ 
and
$\chi^{xxx}_{\mathrm{full-slab}}$,
is very good. Indeed, where the value of $|\chi^{xxx}|$ is large the
difference is very small, and when its value is small, the difference
are just slightly larger, but the spectra is so close to zero that the
differences are not very meaning full; this diferences would become
smaller as the number of atomic layers is increased. However, with 32
layers is more than enough to 
confirm that the extraction of the surface second
harmonic susceptibility from the $2\times 1$ surface is 
readily possible using the formalism contained in Eq.~\eqref{chis}.

Although in a forthcoming publication we will present a study of SSHG
from several Si surfaces where experimental results are available, we
proceed to explain, without confronting with experiment ({\color{red}
Nicolas are there any for this particular component?}), some of the
features seen in 
$|\chi^{xxx}_{2\times 1}|$,  
that is of coures given by
$|\chi^{xxx}_{\mathrm{half-slab}}|$.
First of all, from Fig.~\ref{fig2} 
we see a series of resonances 
that come from the 1$\omega$ and 2$\omega$ terms in
Eq.~\eqref{chis}. The 2$\omega$ resonances naively would start below
$E_g/2$ where $E_g=3.4$ eV is
the band gap of bulk Si. However we see that there are several
resonances well below 1.7 eV, and of course, these resonances are
coming from the electronic states of the $2\times 1$ surface, that lie
inside the gap, that are the well known surface states.
We see that the effect of $\bfv^\nl$ reduces the value of   
$|\chi^{xxx}_{2\times 1}|$,\cite{note2}  
without changing the form of the spectra, as can be confirmed from
both cases of zero and non-zero scissors correction.
Also, we see that the scissors correction shifts the spectra to higher
energies, as expected, however, contrary to the case of linear optics,\cite{cabellosPRB09}
the shifting is not rigid. This is because, the second harmonic optical
response mixes 1$\omega$ and 2$\omega$ transitions, as can be seen from
Eq.~\eqref{chis}, and thus the spectra is not rigidly shifted. 
\begin{figure}
\centering 
\includegraphics[scale=.8]{plots/fig3}
\caption{(color on line) 
$\chi^{xxx}_{2\times 1}$
vs $\hbar\omega$ for a slab of 32 
atomic Si layers plus one layer of H, for two different values of 
the scissors correction $\hbar\sigma$.
\label{fig3}} 
\end{figure}

Finally, in Fig.~\ref{fig3} we compare 
$|\chi^{xxx}_{2\times 1}|$ for two values of $\hbar\sigma$. Indeed,
the value of $\hbar\sigma=0.5$ eV, used above, is the ``average'' GW gap taken from 
Refs.~\onlinecite{rholfingPRB95} which is in agreement with \onlinecite{garciaCPC01}, and  
the value of $\hbar\sigma=0.63$ eV is the ``average'' GW gap taken from 
Refs.~\onlinecite{asahiPRB00}, where more $\bfk$ points in the
Brillouin zone were used.
We see that the fist two peaks are almost rigidly shifted with a small
difference in height, while the rest of the peaks are modified
substantially. This behavior, comes from the fact that the first two
peaks are almost related to the 1$\omega$ resonances of
Eq.~\eqref{chis}, and the other peaks are a combination of 1$\omega$
and 2$\omega$ resonances, thus giving a more different spectrum. This
way we see that small changes in the value of the scissors shift can affect the
$|\chi^{xxx}_{2\times 1}|$ quite dramatically.

@@

However, theoretical development of the field is still an ongoing 
subject of research. Some recent advances for the cases of semiconducting 
and metallic systems have appeared in the literature, where the use of 
theoretical models with experimental results have yielded correct 
physical interpretations for observed SHG spectra.
\cite{
downer_optical_2001,
mendozaPRB01,
lim_optical_2000,
gavrilenko_optical_2000,
mendozaPRB99,
mendozaPRL98,
mendozaPRB96,
mendozaPRB97,
guyotPRB90}

>>>>

\documentclass[floatfix,prb,aps,superscriptaddress,11pt,preprint,letterpaper]{revtex4}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{graphicx}
\allowdisplaybreaks[1]%biutiful equation breaker!!!
\usepackage{ulem}
\usepackage{subfigure}
%\input{/Users/bms/util/definitions}
%\def\chon[1]{\color{red} #1}
\def\chon{red}
\input{definitions}
\usepackage{bm}
%%%%*****************%%%% fine hyperef 
\usepackage[backref,pdffitwindow,colorlinks,linkcolor={blue},citecolor={red}]{hyperref}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{showlabels}
%\usepackage{showkeys}

%%%
% notes:
% nonlocal, not non-local
% check for tense consistency
% surface SHG = SSHG
% $\bfk$-points, not $\bfk$ points
%%%

@@

Concerning the optical response of surfaces and interfaces,
Mendoza and Moch\'an\cite{mendozaPRB96a} produced an early review of 
theoretical SHG spectra from surfaces. 
Arzate and Mendoza\cite{arzatePRB01} improved upon that work and
went into a detailed analysis of the different contributions to the 
SHG spectra of a surface and the nuanced relationship between bulk, 
surface, interband, intraband, $1\omega$ and $2\omega$ terms. 
Calculated surface SHG spectra using these schemes coincide with experimental
results in both the velocity gauge
\cite{mendozaPRB01,mendozaPRL98,reiningPRB94} and the length gauge approach
\cite{sharmaPRB03}. 
Perhaps of more relevance to this work, 
Mendoza \emph{et al}.\cite{mendozaPRB06} and 
Mej\'ia \emph{et al}.\cite{mejiaRMF04} developed a layer-by-layer analysis for 
the linear and nonlinear responses of semiconductor systems. This model
allows for obtaining results from selected regions within a slab or bulk,
including the surface.
More recently, 
Trolle \emph{et al}.\cite{trollePRB14} 
elaborate on a new theoretical model for surface SHG of 
mono-layer MoS$_{2}$. Their approach is based on the tight-binding model 
fitted to GW results and include excitonic effects from the 
Bethe-Salpeter exciton picture. However, this new model is not 
\emph{ab initio} and has no consideration for a layered approach.

@@

In this section we present a 
{\color{\chon} relevant test case} to check the
consistency of {\color{\chon} our approach. We have selected}
a clean Si(001) surface with a $2\times 1$ surface reconstruction.
The slab for such a surface could be chosen to be centrosymmetric 
by {\color{\chon} creating} the front {\color{\chon} and back} 
surfaces with the same $2\times 1$
reconstruction{\color{\chon} . However, we choose to} terminate one 
of the surfaces with hydrogen {\color{\chon} producing an ideal  
terminated} bulk Si surface. {\color{\chon} The} H atoms simply saturate the 
dangling bonds of the bulk-like Si atoms at the surface, as seen in 
Fig.~\ref{si2x1}.
\textcolor{red}{details of the relaxation and some references}
\begin{figure}
\centering 
\includegraphics[scale=.8]{images/si2x1-crop}
\caption{(color on line) The slab shows a 
    {\color{\chon} clean Si(001)$2\times 1$ front surface 
    with an ideal terminated Si bulk back surface. The}
    dangling bonds are H (small balls) saturated. 
    This {\color{\chon} image depicts} 12 Si atomic layers
    {\color{\chon} with one} H atomic layer. 
\label{si2x1}} 
\end{figure} 
The idea {\color{\chon} behind this slab configuration} is that the 
cristalline symmetry of the H terminated surface imposes that 
$\chi_{\mathrm{H}}^{xxx}=0${\color{\chon} . The $2\times 1$ surface has no 
such restrictions, so} $\chi_{2\times 1}^{xxx}\ne 0$.
Thus, calculating $\chi^{xxx}$ for the full-slab, or the 
{\color{\chon} half-slab containing} the $2\times 1$ surface\cite{nota1}
{\color{\chon} should yield the} same result since the contribution from the H
saturated surface is zero {\color{\chon} regardless. 
We must check that the following 
relationship is satisfied for this particular slab}
\begin{align}\label{hs}
\chi_{\mathrm{half-slab}}^{xxx}(-2\go;\go,\go) 
=
\chi_{\mathrm{full-slab}}^{xxx}(-2\go;\go,\go)
,
\end{align}
where
$\chi_{\mathrm{half-slab}}^{xxx}(-2\go;\go,\go)$ is calculated using
$C^\ell(z)=1$ {\color{\chon} from the upper half containing the }$2\times 1$ 
surface reconstruction, as seen in Fig.~\ref{si2x1},
and $\chi_{\mathrm{full-slab}}^{xxx}(-2\go;\go,\go)$ is calculated using
$C^\ell(z)=1$ through the full slab.
{\color{\chon} We show the results for this comparison in the remainder 
of this section}

The self-consistent ground state and their Kohn-Sham states were
calculated in the {\color{\chon} DFT-LDA framework using} the plane-wave 
ABINIT code.\cite{abinit}
We used Troullier-Martins pseudopotentials\cite{troullierPRB91} that are 
fully separable nonlocal pseudopotentials in the Kleinman-Bylander 
{\color{\chon} form.\cite{kleinmanPRL82}
The contribution of $\bfv^\nl$ to Eq.~\eqref{chis} can be carried out 
as detailed in Appendix \ref{vesnl} using the DP code for 
the actual calculation}.\cite{francesco}
The $2\times 1$ and {\color{\chon} H terminated} surfaces are relaxed 
\textcolor{red}{Nicolas some details}, 
{\color{\chon} and we} find agreement with previous results for 
both surfaces.\cite{relax} 
{\color{\chon} For instance the Si-H bond distance is 1.48 \AA ??.} 
The unit cell has two atoms {\color{\chon} making for a computationally
expensive calculation.} 
The vacuum size is equivalent to one {\color{\chon} quarter the} 
size of the slab, {\color{\chon} avoiding the effects produced by possible 
wave-function tunneling} from the contiguous surfaces of the full
crystal formed by the repeated super cell scheme.\cite{mendozaPRB06}

{\color{\chon} Spin-orbit, local field, and electron-hole 
attraction\cite{leitsmannPRB05,trollePRB14} 
effects on the SHG process are all neglected.}
Although these are important factors in the optical response of a semiconductor,
{\color{\chon} their efficient calculation is still numerically challenging and 
under debate. This merits further study but is beyond the scope of this paper.}
For a given {\color{\chon} slab size, we find the} converged spectra 
{\color{\chon} to obtain the relevant parameters. The most important of 
these are: an energy cut-off of 10 Ha, an equal number of conduction and 
valence bands, and a set of 244 $\bfk$-points.} 
{\color{\chon} The $\bfk$-points are used} for the linear analytic 
tetrahedron method {\color{\chon} for evaluating the}
Brillouin zone integrals in Eq.~\eqref{chis}, 
where special care was taken to examine the double resonances.\cite{nastosPRB05}
{\color{\chon}
Double resonances occur at a given frequency $\omega$ if there can be 
resonant transitions at both $\omega$ and 2$\omega$ frequencies. This 
implies that there is a region in the Brillouin zone where 
$\omega_{cv}(\bfk)=2 \omega_{c'v}(\bfk)$. Since there is a real population 
excitation, the perturbation theory used to calculate the spectrum breaks 
down at these $\bfk$-points. This must be taken into account in a more 
complete calculation. These points introduce sharp spikes in the spectrum. 
In principle, these can affect the low frequency results since the response 
is computed from the Kramers-Kronig relation for frequencies below the band 
gap. However, in agreement with Ref.~\onlinecite{nastosPRB05} we find that 
the double resonances affect the low frequency results by less than 2\%. 
All spectra were calculated with a Gaussian smearing of 0.15 eV.}
\begin{figure}
\centering 
\includegraphics[scale=.8]{plots/fig1}
\caption{(color on line) 
$|\chi_{\mathrm{half-slab}}^{xxx}|$ vs $\hbar\omega$ 
{\color{\chon} for the slab} 
with 16, 24, 32, and 40 atomic Si layers. The front surface 
{\color{\chon} is in a clean $2\times 1$ reconstruction and the back 
surface is an ideal terminated bulk H-saturated dangling bonds.}
\label{fig1}} 
\end{figure}
\begin{figure}
\centering 
\includegraphics[scale=.8]{plots/fig2}
\caption{(color on line) 
$\chi^{xxx}_{\mathrm{half-slab}}$
and 
$\chi^{xxx}_{\mathrm{full-slab}}$
vs $\hbar\omega$ for a slab {\color{\chon} with} 32 
atomic Si layers plus one {\color{\chon} H layer}. 
\label{fig2}} 
\end{figure}

{\color{\chon} 
We must evaluate the $[\bfr,\bfv^\nl]$ commutator in order to obtain 
Eqs.~\eqref{tau.1} and \eqref{tau.1n} that are required for 
Eq.~\eqref{chis}, as outlined in Appendix \ref{appvnl}. Computing second 
order derivatives is required thus making the numerical procedure very 
time consuming. This adds significantly to the already lengthy time needed 
for the calculation of the $\bfv^\nl$ contribution and is proportional to 
the first order derivatives. Details can be found in Appendix \ref{vesnl}. 
Memory requirements are also increased for both $\bfv^\nl$ and 
$[\bfr,\bfv^\nl]$. However, the contribution from $[\bfr,\bfv^\nl]$ 
is very small\cite{valerie} and therefore we neglect it in this work.}

In Fig.~\ref{fig1} we show $|\chi_{\mathrm{half-slab}}^{xxx}|$
for {\color{\chon} the} slab {\color{\chon} with} 
16, 24, 32, and 40 Si atomic layers. 
{\color{\chon} We note that the response differs substantially 
for 16 and 24 layers but is quite similar for 32 and 40 layers.}
As explained above,
the calculation of {\color{\chon} the $\bfv^\nl$ contribution is 
computationally expensive.} 
A good compromise between the accuracy in the convergence of
$\chi^{xxx}_{\mathrm{half-slab}}$ as a function of the number
of layers in the slab, and the computational 
{\color{\chon} expense is to consider} 
the slab with 32 Si atomic layers as 
{\color{\chon} an accurate representation} of our 
{\color{\chon} system.}
In Fig.~\ref{fig2}
we compare 
$\chi^{xxx}_{\mathrm{half-slab}}$  
vs. 
$\chi^{xxx}_{\mathrm{full-slab}}$ 
{\color{\chon} for} the four different possibilities 
{\color{\chon} between} including or not including the
effects of $\bfv^\nl$ {\color{\chon} or the} scissors correction
$\hbar\sigma$.
We see that for all four instances the 
{\color{\chon} difference between responses is quite small}. 
{\color{\chon} Indeed, when the value $|\chi^{xxx}|$ 
is large the difference between the two is very small; 
when the value is small the difference increases only slightly, 
but the spectra is so close to zero that these it is negligible. 
These differences would decrease as the number of atomic layers 
increases. We note that 32 layers in the slab} is more than enough 
to confirm that the extraction of the surface second harmonic 
susceptibility from the $2\times 1$ surface is readily possible 
using the formalism contained in Eq.~\eqref{chis}.

{\color{\chon} We will present a study of SSHG from several 
Si surfaces in a forthcoming publication with comparison to experimental 
results. However, without referring to experiment, we proceed to explain 
some of the features seen in $|\chi^{xxx}_{2\times 1}|$ that are obtained 
by calculating $|\chi^{xxx}_{\mathrm{half-slab}}|$.}
({\color{red} Nicolas are there any for this particular component?})
{\color{\chon} First,} from Fig.~\ref{fig2} we note a series of resonances 
that {\color{\chon} derive from} 1$\omega$ and 2$\omega$ terms in
Eq.~\eqref{chis}. {\color{\chon} We could naively assume that the
2$\omega$ resonances} start below $E_g/2$ where $E_g=3.4$ eV is
the band gap of bulk Si. However, we also {\color{\chon} observe} 
that there are several resonances well below 1.7 eV
{\color{\chon} that come} from the electronic states of the 
$2\times 1$ surface, that lie inside the gap and are the 
well known surface states.
We see that the effect of $\bfv^\nl$ reduces the value of   
$|\chi^{xxx}_{2\times 1}|$,\cite{note2}  
without changing the 
{\color{\chon} spectral shape}, as can be confirmed from
both cases of zero and non-zero scissors correction.
{\color{\chon} We note that the scissors correction 
shifts the spectra to higher energies as expected. However, contrary 
to the case of linear optics\cite{cabellosPRB09} this shit is not 
rigid. This is because the second harmonic optical response mixes 
$1\omega$ and $2\omega$ transitions (see Eq.~\eqref{chis}), and accounts 
for the non-rigid shift.}
\begin{figure}
\centering 
\includegraphics[scale=.8]{plots/fig3}
\caption{(color on line) 
$\chi^{xxx}_{2\times 1}$
vs $\hbar\omega$ for a slab {\color{\chon} with} 32 
atomic Si layers plus one {\color{\chon} H layer}, 
for two different values of 
the scissors correction $\hbar\sigma$.
\label{fig3}} 
\end{figure}

Finally, in Fig.~\ref{fig3} we compare 
$|\chi^{xxx}_{2\times 1}|$ for two values of $\hbar\sigma$. 
{\color{\chon} The} value of $\hbar\sigma=0.5$ eV, used above, 
is the ``average'' GW gap taken from 
Refs.~\onlinecite{rholfingPRB95} 
{\color{\chon} and} is in agreement with \onlinecite{garciaCPC01}. 
{\color{\chon} The} value of $\hbar\sigma=0.63$ eV is the ``average'' 
GW gap taken from Refs.~\onlinecite{asahiPRB00}, 
where more {\color{\chon}$\bfk$-points} in the Brillouin zone were used.
We see that the {\color{\chon} first} two peaks are almost rigidly 
shifted with a small difference in {\color{\chon} height while} 
the rest of the peaks are modified substantially. 
This {\color{\chon} behavior comes} from the fact that the first two
peaks are {\color{\chon} almost exclusively related} to the 
1$\omega$ resonances of
Eq.~\eqref{chis}{\color{\chon} . The} other peaks are a combination 
of 1$\omega$ and 2$\omega$ resonances 
{\color{\chon} and yield a more varied} spectrum. 
This way we see that small changes in the value of the 
scissors shift can affect the
$|\chi^{xxx}_{2\times 1}|$ {\color{\chon} spectrum} quite dramatically.

@@

However, theoretical development of the field is still an ongoing 
subject of research. Some recent advances for the cases of semiconducting 
and metallic systems have appeared in the literature, where the use of 
theoretical models with experimental results have yielded correct 
physical interpretations for observed SHG spectra.
\cite{
downer_optical_2001,
mendozaPRB01,
lim_optical_2000,
gavrilenko_optical_2000,
mendozaPRB99,
mendozaPRL98,
mendozaPRB96a,
mendozaPRB97,
guyotPRB90}
